---
title: "Cortexomics"
author: "Dermot Harnett"
date: "July 30, 2018"
output:
  github_document
---



```{r setup, include=FALSE}
#output: rmarkdown::github_document
library(knitr)
```



##Intro
- Importance of understanding gene regulation
- Importance of understanding gene regulation at multiple levels
- Specific importance to the brain
	- complexity of translational regulation in the brain
	- importance of complexes
- Distinction between intra and inter gene variation, discussion of it's importance
- Lack of studies in living tissue that examine Gene expression at multiple levels
-  


###Refs to include


##results
Clustering


### Figure 1 - Data Quality, 
```{r pdf2 nap, out.width ='900px', out.height ='900px' ,fig.cap = "caption", dev='pdf'}
include_graphics("plots/tmp.pdf")
```
Rnaseq, Riboseq, Mass spec correlations, outlier robust correlations.
- Number, size, mapped reads, correlation between replicates
- Enzymes used, for RIboseq, rationale
- Likely effects this has on the data - e.g. cyclohexamine (see sup figure for sequence preference)


### Figure 2 Riboseq
```{r pdf2 nap, out.width ='900px', out.height ='900px' ,fig.cap = "caption", dev='pdf'}
include_graphics("plots/tmp.pdf")

```
Basic discussion of Riboseq
- Periodicity in our Riboseq
- Number of ORFs recovered
- New ORFS recoovered - e.g. 
- Translational Regulation Tests - Ribodiff


### Figure 3 Mass Spec Integration
```{r pdf2 nap, out.width ='900px', out.height ='900px' ,fig.cap = "caption", dev='pdf'}
include_graphics("plots/tmp.pdf")

```
 - Observation that many genes don't have perceptible reaction to changing riboseq levels
 - calculation of degredation rates?
 -  

### Figure 4 - clustering/sequence analysis
```{r pdf2 nap, out.width ='900px', out.height ='900px' ,fig.cap = "caption", dev='pdf'}
include_graphics("plots/tmp.pdf")```

- possible clustering methods
	- timeless
	- 

- Sequence analysis

##Discussion
Resource for study of cortical development
For comparison to single cell data



## Supplementary Figures

###Supp Figure 1

###Supp Figure 2
- Correlation between Riboseq Replicates
- Sequence biases in the Riboseq replicates
- Attemps to compensate?
'We leveraged replicate measurements to assess data quality and its dependence on several parameters, including alignment strategy, mRNA enrichment method, PCR artifacts, gene length normalization, and batch effects (Supplemental Fig. S2A–E; Supplemental Methods).''
- 

###Supp Figure 3

###References
Javonavic et al 2015
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4506746/

Cenik et al 2015
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4617958/
 - Lots of influence on translation by variants in start codons, ORFs.
 - Inter individual variation

Yuzwa et al 2017 cell rep
Developmental Emergence of Adult Neural Stem Cells as Revealed by Single-Cell Transcriptional Profiling.
https://www.ncbi.nlm.nih.gov/pubmed/29281841


Cho et al 2015
memory and translational reg
Methods should be useful, note they measure enrichment of reads in coding sequences.

Zahr et al 2018
https://www.ncbi.nlm.nih.gov/pubmed/29395907



###Methods

