---
title: A-site Alignment (Slide 1)
output:
  ioslides_presentation:
    slide_level: 2
---

```{r setup, echo = FALSE}
  # powerpoint_presentation:
    # slide_level: 2
  # beamer_presentation:
    # slide_level: 2
library(knitr)
library(here)
opts_chunk$set(cache.path=(here::here('Reports/cache/')))

```

# Alignment by A/P site method

## test fig

```{r,include=TRUE, echo=FALSE, fig.caption='foo'}
plot(1)
```

## test fig pdf

:::::::::::::: {.columns}
::: {.column}
```{r, include=TRUE,echo=FALSE,cache=TRUE, out.with='75%', out.height='%75',fig.cap='foo'}
knitr::include_graphics(('sh_fppos_vs_codon_variance.pdf'))
```
:::
::: {.column}
Some text explaining this figure
:::
::::::::::::::

## test fig no cols but pdf

```{r, include=TRUE,echo=FALSE,cache=TRUE, out.with='75%', out.height='%75',fig.cap='foo'}
knitr::include_graphics(('sh_fppos_vs_codon_variance.pdf'))
```

## test fig but direct link

![foo](sh_fppos_vs_codon_variance.pdf)



# Inferred A-site more closely related to Amino acid using Metacodon method (Slide 4)

# Resulting A-site distribution not related to tRNA levels (ab or availability)

- actual availibility of tRNA not necesarily the same as concentration in cell

### Resulting P-site distribution is - at earlier time points

- suggests 

### Also agrees well with poly DT measurements -->
